---
title: "Vegation Recovery Rate Zonification In PR"
author: 
  - "Israel O. Dilán-Pantojas"
  - "Marcelo Francia-Hinostroza"
  - "Manuel J. La Torre-Poueymirou"
date: "April 9, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: 
    highlight: pygments
---

```{r global_options, message=TRUE, warning=TRUE, include=FALSE, paged.print=TRUE}
knitr::opts_chunk$set(echo = T)
```
## Methods: 

## References:


\pagebreak

## Apendix:
```{r loading_libraries, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ncdf4) # package for netcdf manipulation
library(raster) # package for raster manipulation
library(rgdal) # package for geospatial analysis
library(rasterVis)
library(RColorBrewer)

# library(maptools)
# library(maps)
# library(MODISTools)
# library(mapdata)
library(ggplot2) # packages for plotting
library(chron)
# library(lattice)

```

```{r loading_data}
# load data
setwd("~/Desktop/BigData2Biol/Project_2/Hot_zones")
data <- nc_open(filename = "MOD13Q1.006_250m_aid0001.nc")

# get longitude and latitude
lon <- ncvar_get(data, "lon")
nlon <- dim(lon)
lat <- ncvar_get(data, "lat")
nlat <- dim(lat)

# get time dimension
time <- ncvar_get(data, "time")
tunits <- ncatt_get(data,"time","units")
nt <- dim(time)

# get ndvi
dname <- "_250m_16_days_NDVI"
ndvi_array <- ncvar_get(data,dname)
dlname <- ncatt_get(data,dname,"long_name")
dunits <- ncatt_get(data,dname,"units")
fillvalue <- ncatt_get(data,dname,"_FillValue")

# get LST
# dname <- "_LST"
# ndvi_array <- ncvar_get(data,dname)
# dlname <- ncatt_get(data,dname,"long_name")
# dunits <- ncatt_get(data,dname,"units")
# fillvalue <- ncatt_get(data,dname,"_FillValue")

# get evapo
# dname <- "_EVAP"
# ndvi_array <- ncvar_get(data,dname)
# dlname <- ncatt_get(data,dname,"long_name")
# dunits <- ncatt_get(data,dname,"units")
# fillvalue <- ncatt_get(data,dname,"_FillValue")

# get global attributes
title <- ncatt_get(data,0,"title")
institution <- ncatt_get(data,0,"institution")
datasource <- ncatt_get(data,0,"source")
references <- ncatt_get(data,0,"references")
history <- ncatt_get(data,0,"history")
Conventions <- ncatt_get(data,0,"Conventions")

nc_close(data)

# convert time -- split the time units string into fields
tustr <- strsplit(tunits$value, " ")
tdstr <- strsplit(unlist(tustr)[3], "-")
tmonth <- as.integer(unlist(tdstr)[2])
tday <- as.integer(unlist(tdstr)[3])
tyear <- as.integer(unlist(tdstr)[1])

# replace netCDF fill values with NA's
# ndvi_array[ndvi_array==fillvalue$value] <- NA
NDVIS <- ndvi_array

# identify selected dates
dates <- chron(time,origin=c(tmonth, tday, tyear))
```

<!-- Possible seasonality:  https://verbe039.github.io/BFASTforAEO/ -->

```{r Analysis}
# https://www.earthdatascience.org/courses/earth-analytics/multispectral-remote-sensing-modis/normalized-burn-index-dNBR/
# 
# library(ggplot2)
# library(ggmap)

c <- raster('MOD13Q1.006_250m_aid0001.nc')
s <- stack(c)
plot(s)

# use colorbrewer which loads with the rasterVis package to generate
# a color ramp of yellow to green
cols <- colorRampPalette(brewer.pal(9,"YlGn"))
# create a level plot - plot

levelplot(s, main="Landsat NDVI -- Improved Colors \nNEON Harvard Forest Field Site", col.regions=cols)


# https://www.neonscience.org/raster-time-series

# plot NDVI
# avg_NDVI_HARV <- as.data.frame(cellStats(s,mean))
# 
# names(avg_NDVI_HARV) <- "meanNDVI"
# avg_NDVI_HARV$site <- "Yunque"
# 
# julianDays <- gsub("X|_HARV_NDVI_crop", "", row.names(avg_NDVI_HARV))
# 
# avg_NDVI_HARV$julianDay <- julianDays
# 
# ggplot(avg_NDVI_HARV, aes(julianDay, meanNDVI), na.rm=TRUE) +
#   geom_point(size=4,colour = "PeachPuff4") + 
#   ggtitle("Landsat Derived NDVI - 2011\n NEON Harvard Forest Field Site") +
#   xlab("Julian Days") + ylab("Mean NDVI") +
#   theme(text = element_text(size=20))
# 
# # map the data
# world.outlines <- map("world" , plot=T)
# world.outlines.sp <- map2SpatialLines(world.outlines, proj4string = CRS("+proj=longlat"))
# 
# mapTheme <- rasterTheme(region = rev(brewer.pal(10, "RdBu")))
# 
# yunque <- (ggmap(get_map(location = c(nlon,nlat))))
# yunque + plot(s)
# 
#  
# plt <- levelplot(s, main="Landsat NDVI -- Improved Colors \nNEON Harvard Forest Field Site")
# plt + layer(ggmap(get_map(location = c(nlon,nlat))))

png("zonification_plot.png")
# Stack plot for zones
levelplot(s, main="Landsat NDVI stacked over time", col.regions = colorRampPalette(brewer.pal(10,"RdYlGn")))
dev.off()

# Stack plot for temp
levelplot(s, main="Landsat LST stacked over time", col.regions = colorRampPalette(brewer.pal(10,"RdYlBu")))

# Stack plot for evap
levelplot(s, main="Landsat Evapo stacked over time", col.regions = colorRampPalette(brewer.pal(10,"BrBG")))
```

```{r}
for (m in 1:length(dates)) {
  
  par(mfrow=c(1,3))
  ndvi_slice <- ndvi_array[,,m]
  qual_slice <- qual_array[,,m]
  
  r <- raster(t(ndvi_slice), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
  plot(r, main = dates[m] )
    # turn off scientific notation
  options("scipen" = 100, "digits" = 4)
  # bottom, left, top and right
  #par(mfrow=c(4, 2))
  hist(ndvi_slice,
    col = "springgreen",
    xlab = "NDVI Value")
  mtext(paste("Distribution of MODIS NDVI ", dates[m]),
        outer = TRUE, cex = 1.5)
  hist(qual_slice,
    col = "skyblue",
    xlab = "Quality Value")
  mtext(paste("Distribution of MODIS Quality ", dates[m]),
        outer = TRUE, cex = 1.5)
}
# dev.copy(png,'GIMMS3g_1982_NDVI.png')
# dev.off()
```

First, we will need to convert the entire 3d array of data to a raster brick. **Note, this step may take several minutes.**
```{r rasterize_brick}
r_brick <- brick(ndvi_array, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
# note that you may have to play around with the transpose (the t() function) and flip() before the data are oriented correctly. In this example, the netcdf file recorded latitude on the X and longitude on the Y, so both a transpose and a flip in the y direction were required.
# r_brick <- flip(t(r_brick), direction='y')
```

Extract timeseries of data at the Toolik Lake study location from the raster brick using the 'extract()' function.
```{r get_Toolik}
# El yunque-ish 18.2724609375, -65.80810546875
toolik_lon <- 18.2724609375
toolik_lat <- -65.80810546875
toolik_series <- extract(r_brick, SpatialPoints(cbind(toolik_lon,toolik_lat)), method ="simple")
```

This timeseries is in a simple vector indexed only by the raster layer ID, so let's put it in an easier-to-use dataframe form and then plot the timeseries.
```{r plot_Toolik}
toolik_df <- data.frame(year=dates, NDVI=t(toolik_series)) # save as a dataframe
# plot and save
ggplot(data=toolik_df, aes(x=year, y=NDVI, group=1)) +
  geom_line() + # make this a line plot
  ggtitle("Growing season NDVI at Toolik Lake Station") +     # Set title
  theme_bw() # use the black and white theme
# ggsave('GIMMS3g_Toolik.png')
```
